---
title: "Lab 6"
author: "Katherine Rosencrance"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)

```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t test)

H0: equal ranks or medians
HA: Ranks are not equal or medians are not equal

### Create samples

`set.seed()` to create pseudorandom but reproducible samples

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

gp_1

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)

gp_2
```
Take a quick look at distributions using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t test so we can compare results

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument paired = TRUE within t test
```

This p-value means that if our samples are drawn from populations with the same mean (ie if the null hypothesis is true) there is `r round(ttest$p.value,2)` probability of finding sample means *at least as different* as mine by random chance

## Mann- Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

this p value means that if our samples are drawn from populations with the same median the probability of finding sample median at least as difference as mine by random chance is `r round(wilcox.test$p.value,2)`

no significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round( mwu$p.value,2)`)

## Simple linear regression

two *continuous quantitative* variables that you are trying to understand the relationship between and you think that a linear relationship makes sense


Use penguins data from palmer penguins exploring flipper length vs body mass

```{r}
# Make an exploratory scatter plot of flipper length vs body mass 

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g))+
  geom_point() +
  geom_smooth(method = "lm")
```
### FInd the model estimates for coefficients

Use the `lm()` function for linear model

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

model coefficients: `r penguin_lm$coefficients`
fliiper length coefficient: `r penguin_lm$coefficients[2]`

### use broom::tody to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```


### use broom::glance to ge output of model fit and staistical significance

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```


### USe equatiomatic::extract_eq to return the model in LateX

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```
### use plot() to automatically produce diagnostic plots

```{r}
plot(penguin_lm)
```

